Tree height and hydraulic traits shape growth responses across droughts in a temperate broadleaf forest

### 2020-07-13 manuscript/response_to_reviews.Rmd

**We switched to use of generalized linear mixed models. This did not change the outcome of any analyses.**

[*issue #94*](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/94)

I suggest to show some relationship, e.g. between Rt and traits in figures. The study only presents two figures and there is plenty of room to present results more clearly with extra figures.

**We created a new figure (*Fig. 4*) visualizing the top model results.**  ([*issue #71*](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/71)()

@@

[*issue #94*](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/94)

**We have gone through the manuscript / tables and figures and updated where it said "dAICc" to be $\Delta$AICc. **

Following the two previous comments, I suggest to leave in Table 5 just the best model (rather that a multimodel inference as is presented now) to discuss more clearly the covariates included. And I suggest discussing only the model with all years together. This would be more robust and avoid ramble differences between years that mostly look spurious and with a difficult physiological explanation. This would help to ensure a better use of the Resilience indices.

**In the spirit of this comment (simplifying presentation), we have created a new figure (*Fig. 4*) visualizing the top model results. We moved Table 5 to the Appendix (now Table S6, plus a parallel Table S7 for $Rt_{ARIMA}$).**

@@

**We do want to show the three separate drought periods, as we feel that this makes the analysis richer and helps to separate real drivers from spurious correlations. However, in order to make models more consistent across years and reduce spurious correlations, we now include the criteria that for a trait to be included as a candidate variable in the full model, the sign of coefficients must be consistent across all droughts (in addition to meeting the AIC criteria, as before). This resulted in only two candidate variables for the main model ($PLA_{dry}$ and $\pi_{tlp}$). (Removed was ring porosity, which had opposite effects in different droughts.)**

The authors might like to consider including Table 4 as a supplement, because as it stands it looks like an exploratory analysis. If the ‘single-variable approach’ is not just exploratory, state it clearly and why it is needed besides the full models with several parameters.

**We have moved this table to the SI. We modified the analysis approach such that separate models for individual variable tests (as previously show in this table) are now limited to the species traits. While this table is (and was) the only display item (besides Table 1) giving these results, the results are either mostly negative (wood density, SLA, xylem porosity) or preliminary tests qualifying variables for inclusion in full model ($PLA_{dry}$ and $\pi_{tlp}$).**

@@

**It is true that crown position estimates become more accurate in later dates, but height should be equally reliable across years, having been estimated based on reconstructed DBH. This was not previously clearly specified in the methods. To clarify, we have added the following statement: "We then used these allometries to estimate $H$ for each drought year, $Y$, based on reconstructed $DBH_Y$."**

**We agree that the assessment of crown position is somewhat problematic, both because of its increasing inaccuracy going back through time, and because of collinearity with $H$. However, given that it is both biologically interesting and comes out in our top models, we have chosen to include it. We note that this makes the test of our hypothesis conservative. We are transparent about this issue, to which we devote a paragraph in the discussion (par. 3).**

- Line 275-280: those profiles were assumed a priori in the introduction and are common to any forest. They are indeed interesting to characterise the forest, but they do not add to individual tree variability. You could think to send Figure 2 to an appendix, in consequence, and use figures to show more specific results related to your individual tree analyses.

[*issue #102*](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/10)

**We did create two new figures showing the results of the tree-ring analysis (and moved two tables to the SI).**

@@

**Thank you.**

105 this sounds great, but was a little unclear. When you say “first, we test…” I thought that you would examine a relationship between microsite and some tree characteristic, e.g. that the largest trees are in moist microsites. But the next sentence didn’t refer to a tree-microsite correlation but vulnerability-size. It would be good to clarify the prose here because these are very important passages.  The third sentence did seem to be about tree-microsite correlations.

112 wasn’t clear where the “ring porous more drought resistant” hypothesis came from. Global data show that ring porous species are almost entirely restricted to the north temperate zone, especially deciduous broadleaf forests (Wheeler et al., 2007). They are virtually absent in drylands, especially warm tropical drylands. Species in drylands are almost overwhelmingly diffuse porous; the high wood density species in chaparrals and in some desert plants can be semi ring porous (Olson et al., in press).

*good point. hypothesis refers to temperate deciduous biome, and is not supported here.*

### 2020-07-13 manuscript/response_to_reviews.Rmd

**It is true that crown position estimates become more accurate in later dates, but height should be equally reliable across years, having been estimated based on reconstructed DBH. This was not previously clearly specified in the methods. To clarify, we have added the following statement: "We then used these allometries to estimate $H$ for each drought year, $Y$, based on reconstructed $DBH_Y$."**

**We agree that the assessment of crown position is somewhat problematic, both because of its increasing inaccuracy going back through time, and because of collinearity with $H$. However, given that it is both biologically interesting and comes out in our top models, we have chosen to include it. We note that this makes the test of our hypothesis conservative. We are transparent about this issue, to which we devote a paragraph in the discussion (par. 3).**

- Line 275-280: those profiles were assumed a priori in the introduction and are common to any forest. They are indeed interesting to characterise the forest, but they do not add to individual tree variability. You could think to send Figure 2 to an appendix, in consequence, and use figures to show more specific results related to your individual tree analyses.

**Our preference is to keep this figure in the main manuscript. We agree that similar patterns are common to any forest, and are not trying to claim this as any kind of novel result, but we're not aware of studies showing similar height profiles for a temperate forest. Such patterns are not shown some relevant textbooks: Gordan Bonan's 2016 *Ecological Climatology* and Norman & Campbell's *Environmental Biophysics*. If we're missing studies that show such vertical profiles, we'd appreciate the references.**

### 2020-07-13 manuscript/tables&figures.Rmd
added response on keeping vertical profiles fig in main doc
closes #102
update tables & figures

Parentheses indicate that the prediction was supported by one but not all of the top full models (Table **5**). Dash symbols indicate that the response was not significant/ not represented in any of the top full models.
Asterisks indicate that the $Rt_{ARIMA}$ model yielded a different result: 
*different significance in $Rt_{ARIMA}$ model
**$Rt_{ARIMA}$ model had a significant trend in the opposite direction of the $Rt$ model (if significant) or the prediction (if no signficant trend in $Rt$ model).

**TILDAS (~) indicate that we don't currently have a (full) test of this prediction ([issue 99](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/99)**)

@@

![**Figure 4. Visualization of top models for all droughts combined and for each drought separately.** Model coefficients are given in Table S6. **THIS IS NOT YET CORRECT--CANOPY POSITION NEEDS TO BE CHANGED TO CATEGORICAL VARIABLE, AND THE POSITIVE HEIGHT EFFECT FOR 1999 IS ALSO WRONG (PROBABLY ARTIFACT OF CP). FINAL FIG WILL MATCH MODELS IN TABLE S6.** ](tables_figures/publication/regr_visuals.png){width=500px}

>>>>

Parentheses indicate that the prediction was supported by one but not all of the top full models (Table **5**). Dash symbols indicate that the response was not significant/ not represented in any of the top full models.
Asterisks indicate that the $Rt_{ARIMA}$ model yielded a different result: 
*$Rt_{ARIMA}$ had a difference in significance/ inclusion in top full models.
**$Rt_{ARIMA}$ model had a significant trend in the opposite direction of the $Rt$ model (if significant) or the prediction (if no significant trend in $Rt$ model).

@@

![**Figure 4. Visualization of top models for all droughts combined and for each drought separately.** Model coefficients are given in Table S6. **THIS IS NOT YET CORRECT (having trouble making gg-plot work. THE POSITIVE HEIGHT EFFECT FOR 1999 IS WRONG. FINAL FIG WILL MATCH MODELS IN TABLE S6.** ](tables_figures/publication/regr_visuals.png){width=500px}
